---
title: "Religious Attendance and CVD"
output: html_notebook
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(kableExtra)
library(gtsummary)
library(foreign)
#library(texreg)
#library(arm)
```

```{r}
#demo
xpt <- foreign::read.xport("../Data/DEMO_E.XPT")

data <- xpt[c("SEQN", "RIAGENDR", "RIDAGEYR", "RIDRETH1", "DMDEDUC2", "DMDMARTL", "DMDBORN2")] %>% 
  rename(gender = RIAGENDR,
         age = RIDAGEYR,
         ethnicity = RIDRETH1,
         education = DMDEDUC2,
         marital_status = DMDMARTL,
         born_outside_us = DMDBORN2)

#INQ
xpt <- foreign::read.xport("../Data/INQ_E.xpt")

data <- merge(data, xpt[c("SEQN", "INDFMMPI", "INDFMMPC")], by = "SEQN", all.x = TRUE) %>% 
  rename(poverty_level_index = INDFMMPI,
         poverty_level_category = INDFMMPC)

#HIQ
xpt <- foreign::read.xport("../Data/HIQ_E.xpt")

data <- merge(data, xpt[c("SEQN", "HIQ011", "HIQ031A", "HIQ031B", "HIQ031D")], by = "SEQN", all.x = TRUE) %>% 
  rename(health_insurance = HIQ011,
         private_insurance = HIQ031A,
         medicare = HIQ031B,
         medicaid = HIQ031D)

#HUQ
xpt <- foreign::read.xport("../Data/HUQ_E.xpt")

data <- merge(data, xpt[c("SEQN", "HUQ050")], by = "SEQN", all.x = TRUE) %>% 
  rename(num_healthcare_visits = HUQ050)

#FSQ
xpt <- foreign::read.xport("../Data/FSQ_E.xpt")

data <- merge(data, xpt[c("SEQN", "FSQ165", "FSQ171", "FSD670ZC", "FSD032A", "FSD032B")], by = "SEQN", all.x = TRUE) %>% 
  rename(ever_received_food_stamps = FSQ165,
         #food_stamps_past_yr = FSQ171,
         num_months_WIC = FSD670ZC,
         worried_house_run_out_food = FSD032A,
         food_ran_out = FSD032B)

#PFQ
xpt <- foreign::read.xport("../Data/PFQ_E.xpt")

#risk_mobility
data <- merge(data, xpt[c("SEQN", "PFQ049")], by = "SEQN", all.x = TRUE) %>% 
  dplyr::rename(mobility_limit = PFQ049) %>% 
  dplyr::mutate(risk_mobility = case_when(
                mobility_limit == "Yes" ~ 1,
                TRUE ~ as.double(0)))

#BPQ
xpt <- foreign::read.xport("../Data/BPQ_E.xpt")

#risk_bp
data <- merge(data,xpt[c("SEQN", "BPQ020")], by = "SEQN", all.x = TRUE) %>% 
  rename(high_blood_pressure = BPQ020) %>% 
  mutate(risk_bp = case_when(
    high_blood_pressure == "Yes" ~ 1,
    TRUE ~ as.double(0)
  ))


#GLU
xpt <- foreign::read.xport("../Data/GLU_E.xpt")

data <- merge(data, xpt[c("SEQN", "LBDGLUSI")], by = "SEQN", all.x = TRUE) %>%
  rename(fasting_glucose = LBDGLUSI)

#fasting glucose
data <- data %>% 
  mutate(fasting_glucose = case_when(
    fasting_glucose <= 4.995 ~ "Not at Risk",
    fasting_glucose > 4.995 ~ "At Risk",
    is.na(fasting_glucose) ~ "Unknown"
  ))

#risk_glucose
data$fasting_glucose <- factor(data$fasting_glucose, levels = c("Not at Risk", "At Risk"), c("Not at Risk", "At Risk"))
data <- data %>% 
  mutate(risk_glucose = case_when(
    fasting_glucose == "At Risk" ~ 1,
    TRUE ~ as.double(0)
  ))


#SMQRTU
xpt <- foreign::read.xport("../Data/SMQRTU_E.xpt")

data <- merge(data, xpt[c("SEQN", "SMQ680")], by = "SEQN", all.x = TRUE) %>% 
  rename(tobacco_use_5_days = SMQ680)

#ALQ
xpt <- foreign::read.xport("../Data/ALQ_E.xpt")

data <- merge(data, xpt[c("SEQN", "ALQ101", "ALQ130", "ALQ120Q", "ALQ120U", "ALQ140Q", "ALQ140U", "ALQ110")], by = "SEQN", all.x = TRUE) %>% 
  rename(had_12_drinks_in_1yr = ALQ101,
         avg_drinks_daily_past_yr = ALQ130,
         num_drinks_past_yr = ALQ120Q,
         num_drinks_past_yr_unit = ALQ120U,
         num_days_5_drinks_or_more = ALQ140Q,
         had_12_drinks_lifetime = ALQ110)

data <- data %>% 
  mutate(num_drinks_weekly = case_when(
    num_drinks_past_yr_unit == 1 ~ num_drinks_past_yr,
    num_drinks_past_yr_unit == 2 ~ num_drinks_past_yr/4.25,
    num_drinks_past_yr_unit == 3 ~ num_drinks_past_yr/52,
 #   num_drinks_past_yr == 777 ~ NA,
 #   num_drinks_past_yr == 999 ~ NA,
    TRUE ~ as.double(num_drinks_past_yr)
  ))

data <- data %>% 
  mutate(num_days_5_drinks_or_more = case_when(
    #convert to days per week
    ALQ140U == 1 ~ num_days_5_drinks_or_more,
    ALQ140U == 2 ~ num_days_5_drinks_or_more/4.25,
    ALQ140U == 3 ~ num_days_5_drinks_or_more/52,
    TRUE ~ as.double(num_days_5_drinks_or_more)
  ))

data <- data %>% 
  mutate(type_of_drinker = case_when(
    had_12_drinks_lifetime == 2 ~ "Abstainer",
    num_days_5_drinks_or_more >= 1 ~ "Harmful",
    age >= 65 & num_drinks_weekly > 7 ~ "Harmful",
    had_12_drinks_in_1yr == 1 & (age < 65 | num_days_5_drinks_or_more < 1) & gender == 2 & num_drinks_weekly <= 7 ~ "Moderate",
    had_12_drinks_in_1yr == 1 & (age < 65 | num_days_5_drinks_or_more < 1) & gender == 2 & num_drinks_weekly > 7 ~ "Harmful",
    had_12_drinks_in_1yr == 1 & (age < 65 | num_days_5_drinks_or_more < 1) & gender == 1 & num_drinks_weekly > 14 ~ "Harmful",
    had_12_drinks_in_1yr == 1 & (age < 65 | num_days_5_drinks_or_more < 1) & gender == 1 & num_drinks_weekly <= 14 ~ "Moderate",
    TRUE ~ as.character("NA")
  )) %>% 
  mutate(risk_alcohol = case_when(
    type_of_drinker == "Harmful" ~ 1,
    TRUE ~ as.double(0)
  ))

data$type_of_drinker <- factor(data$type_of_drinker, c("Abstainer", "Harmful", "Moderate"), c("Abstainer", "Harmful", "Moderate"))



#SSQ
xpt <- foreign::read.xport("../Data/SSQ_E.xpt")

data <- merge(data, xpt[c("SEQN", "SSD044")], by = "SEQN", all.x = TRUE) %>% 
  rename(religious_attendance = SSD044)

#MCQ
xpt <- foreign::read.xport("../Data/MCQ_E.xpt")

data <- merge(data, xpt[c("SEQN", "MCQ160E", "MCQ160F")], by = "SEQN", all.x = TRUE) %>% 
  rename(heart_attack = MCQ160E,
         stroke = MCQ160F)

#DR1FF
#xpt <- foreign::read.xport("../Data/DR1IFF_E.xpt") %>% 
#  dplyr::select(SEQN, DR1IKCAL) %>% 
#  group_by(SEQN) %>% 
#  summarise(SEQN = SEQN,
#            calories = mean(DR1IKCAL * 1000, na.rm = T)) %>% 
#  distinct()

#data <- merge(data, xpt, by = "SEQN", all.x = TRUE)


data$gender <- factor(data$gender,levels = c(1,2), labels = c("Male", "Female"))

data <- data %>% 
  mutate(ethnicity = case_when(
    ethnicity %in% c(1,2) ~ 1,
    TRUE ~ as.double(ethnicity)
  ))

data$ethnicity <- factor(data$ethnicity, levels = c(3,1,4,5), labels = c( "Non-Hispanic White", "Hispanic", "Non-Hispanic Black", "Other"))

data$education <- factor(data$education, levels = c(1,2,3,4,5), labels = c("Less than 9th grade", "9-11th Grade", "High School/GED", "Some college", "College Graduate or above"))

data <- data %>% 
  mutate(education = case_when(
    education %in% c("Less than 9th grade", "9-11th Grade") ~ "Less than High School",
    TRUE ~ as.character(education)
  ))

data$education <- factor(data$education, levels = c("High School/GED", "Less than High School", "Some college", "College Graduate or above"), labels = c("High School/GED", "Less than High School", "Some college", "College Graduate or above"))

data$marital_status <- factor(data$marital_status, levels = c(1,2,3,4,5,6), labels = c("Married", "Widowed", "Divorced", "Separated", "Never Married", "Living with partner"))

data <- data %>% 
  mutate(marital_status = case_when(
    marital_status %in% c("Married", "Living with partner") ~ "Married",
    marital_status %in% c("Divorced", "Widowed", "Separated") ~ "Previously Married",
    TRUE ~ as.character(marital_status)
  ))

data$marital_status <- factor(data$marital_status, levels = c("Never Married", "Married", "Previously Married", "Refused", "Don't Know"),
                              labels = c("Never Married", "Married", "Previously Married", "Refused", "Don't Know"))


data$born_outside_us <- factor(data$born_outside_us, levels = c(1,2,4,5), labels = c("Born in US", "Born in Mexico", "Born in other spanish speaking country", "Born in other non-spanish speaking country"))

data <- data %>% 
  mutate(born_outside_us = case_when(
    born_outside_us %in% c("Born in Mexico", "Born in other spanish speaking country", "Born in other non-spanish speaking country") ~ " Born outside US",
    TRUE ~ as.character(born_outside_us)
  ))

data$born_outside_us <- factor(data$born_outside_us, levels = c("Born in US", "Born outside US"), labels = c("Born in US", "Born outside US"))


data$poverty_level_category <- factor(data$poverty_level_category, c(1,2,3), c("Poverty Index <= 1.3", "Poverty Index in (1.30, 1.85)", "Poverty Index > 1.85"))

data <- data %>% 
  mutate(health_insurance = case_when(
    is.na(health_insurance) ~ "Unknown",
    TRUE ~ as.character(health_insurance)
  ))

data <- data %>% 
  mutate(private_insurance = case_when(
    is.na(private_insurance) ~ "Unknown",
    TRUE ~ as.character(private_insurance)
  ))

data <- data %>% 
  mutate(medicare = case_when(
    is.na(medicare) ~ "Unknown",
    TRUE ~ as.character(medicare)
  ))

data <- data %>% 
  mutate(medicaid = case_when(
    is.na(medicaid) ~ "Unknown",
    TRUE ~ as.character(medicaid)
  ))

data$health_insurance <- factor(data$health_insurance, c("2","1"), c("Not Covered", "Covered"))
data$private_insurance <- factor(data$private_insurance, c("Unknown", "14"), c("Not Covered", "Covered"))
data$medicare <- factor(data$medicare, c( "Unknown","15"), c("Not Covered", "Covered"))
data$medicaid <- factor(data$medicaid, c( "Unknown", "17"), c("Not Covered", "Covered"))

data$num_healthcare_visits <- factor(data$num_healthcare_visits, c(0,1,2,3,4,5), c("None", "1", "2 to 3", "4 to 9", "10 to 12", "13 or more"))
data$ever_received_food_stamps <- factor(data$ever_received_food_stamps, c(1,2), c("Yes", "No"))
#data$food_stamps_past_yr <- factor(data$food_stamps_past_yr, c(1,2,7,9), c("Yes", "No", "Refused", "Don't Know"))
#data$num_months_WIC targeted 0-11 years
data$worried_house_run_out_food <- factor(data$worried_house_run_out_food, c(1,2,3), c("Often True", "Sometimes true", "Never true"))
data$food_ran_out <- factor(data$food_ran_out, c(1,2,3), c("Often True", "Sometimes true", "Never true"))
data$mobility_limit <- factor(data$mobility_limit, c(1,2), c("Yes", "No"))
data$high_blood_pressure <- factor(data$high_blood_pressure, c(1,2), c("Yes", "No"))


#risk_tobacco
data <- data %>% 
  mutate(tobacco_use_5_days = case_when(
    tobacco_use_5_days == 1 ~ "Yes",
    tobacco_use_5_days == 2 ~ "No",
    tobacco_use_5_days %in% c(7,9) ~ "Unknown",
    is.na(tobacco_use_5_days) ~ "Unknown"
  )) %>% 
  mutate(risk_tobacaco = case_when(
    tobacco_use_5_days == "Yes" ~ 1,
    TRUE ~ as.double(0)
  ))
data$tobacco_use_5_days <- factor(data$tobacco_use_5_days, c("Yes", "No"), c("Yes", "No"))


data$had_12_drinks_in_1yr <- factor(data$had_12_drinks_in_1yr, c(1,2), c("Yes", "No"))

data$heart_attack_num <- data$heart_attack
data$stroke_num <- data$stroke

data$heart_attack <- factor(data$heart_attack, c(1,2), c("Yes", "No"))
data$stroke <- factor(data$stroke, c(1,2), c("Yes", "No"))

#Religious Attendance
data <- data %>% 
  mutate(religious_attendance_num = religious_attendance,
          religious_attendance = case_when(
          religious_attendance < 50 ~ "Less than Weekly",
          religious_attendance %in% c(50,51,52) ~ "Weekly",
          religious_attendance > 52 & religious_attendance < 7777 ~ "More than   Weekly",
          religious_attendance >= 7777 ~ "NA",
          TRUE ~ as.character(religious_attendance)
        ))

data$religious_attendance <- factor(data$religious_attendance, levels = c("Less than Weekly", "Weekly", "More than Weekly"), labels = c("Less than Weekly", "Weekly", "More than Weekly"))




#BMX
xpt <- foreign::read.xport("../Data/BMX_E.xpt")
data <- merge(data, xpt[c("SEQN", "BMXWT", "BMXHT", "BMXWAIST")], by = "SEQN", all.x = TRUE) %>% 
  rename(height = BMXHT,
         weight = BMXWT,
         waist = BMXWAIST)

data <- data %>% 
  mutate(rfm = case_when(
    gender == "Male" ~ as.double(64 - (20 * (height / waist))),
    gender == "Female" ~ as.double(76 - (20 * (height / waist)))
  ))

data <- data %>% 
  mutate(risk_rfm = case_when(
    gender == "Male" & rfm >= 30 ~ 1,
    gender == "Female" & rfm >= 40 ~ 1,
    TRUE ~ as.numeric(0)
  ))

data <- data %>% 
  group_by(SEQN) %>% 
  mutate(risk_sum = sum(risk_glucose, risk_mobility, risk_bp, risk_tobacaco, risk_alcohol, risk_rfm))


data$risk_sum <- factor(data$risk_sum)

data <- data %>% 
  filter(age >= 20)

write.foreign(df=data, datafile="../Data/nhanes0708.csv", codefile="../Data/nhanes0708.sas", package="SAS")
write.csv(data, "../Data/nhanes0708.csv") 
```

# Summary Tables

```{r, warning=FALSE, message=FALSE, warn = FALSE}
library(vtable)

data2 <- data %>% 
  dplyr::select(SEQN, age, poverty_level_index, religious_attendance)

sumtable(data2, out = "kable", summ = c('notNa(x)','(notNa(x)/sum(x,na.rm = TRUE))', 'mean(x)') , summ.names = c("N", 'Percent', "Mean")) %>% 
  kable_styling("striped")

data2 <- data %>% 
  dplyr::select(-c(age, poverty_level_index, religious_attendance))

sumtable(data2, out = "kable", summ = c('notNa(x)','notNa(x)/sum(x,na.rm = TRUE)','mean(x)'), summ.names = c("N", 'Percent', "Mean")) %>% 
  kable_styling("striped")
```

# Logistic Regression Models


### Heart attack and stroke models with all variables

```{r}
#heart attack
heart_model <- glm(heart_attack ~ gender + age + rfm + ethnicity + high_blood_pressure + religious_attendance + education + marital_status 
                     + poverty_level_category + num_healthcare_visits + ever_received_food_stamps + worried_house_run_out_food + food_ran_out +
                     mobility_limit + fasting_glucose + tobacco_use_5_days + type_of_drinker, data = data, family = "binomial")

summary(heart_model)

#Stroke
stroke_model <- glm(stroke ~ gender + age + rfm + ethnicity + high_blood_pressure + religious_attendance + education + marital_status +  poverty_level_category + num_healthcare_visits + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + mobility_limit + fasting_glucose + tobacco_use_5_days + type_of_drinker, data = data, family = "binomial")

summary(stroke_model)
```

### Heart attack and stroke models using social demographics

```{r}
#Social determinant model 
social_heart <- glm(heart_attack ~ gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid, data = data, family = "binomial")

summary(social_heart)


social_stroke <- glm(stroke ~gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid, data = data, family = "binomial")

summary(social_stroke)

#Health/behavior

#religious
```

### Heart attack and stroke models using health variables

```{r}
health_heart <- glm(heart_attack ~ num_healthcare_visits + risk_sum, data = data, family = "binomial")
summary(health_heart)

health_stroke <- glm(stroke ~ num_healthcare_visits + risk_sum, data = data, family = "binomial")
summary(health_stroke)
```

### Heart attack and stroke models using religious attendance alone

```{r}
relig_heart <- glm(heart_attack ~ religious_attendance, data = data, family = "binomial")
summary(relig_heart)

relig_stroke <- glm(stroke ~religious_attendance, data = data, family = "binomial")
summary(relig_stroke)
```

```{r, warn = FALSE, warning = FALSE}
#make tables
#library(broom)
#library(parameters)
social_heart_tbl <- tbl_regression(social_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)  

social_stroke_tbl <- tbl_regression(social_stroke, exponentiate = TRUE) %>%
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

health_heart_tbl <- tbl_regression(health_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

health_stroke_tbl <- tbl_regression(health_stroke, exponentiate = TRUE) %>%
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

relig_heart_tbl <- tbl_regression(relig_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

relig_stroke_tbl <- tbl_regression(relig_stroke, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)
```

### Heart Attack models

```{r}
tbl_merge(tbls = list(social_heart_tbl, health_heart_tbl, relig_heart_tbl),
          tab_spanner = c("Social Determinants", "Health and Behavior", "Religious Attendance"))
```


### Stroke models

```{r}
tbl_merge(tbls = list(social_stroke_tbl, health_stroke_tbl, relig_stroke_tbl),
          tab_spanner = c("Social Determinants", "Health and Behavior", "Religious Attendance"))
```

# Step Wise Logistic Regression

### Social 
```{r}
#social data
social <- data %>% 
  select(heart_attack, stroke, gender, age, ethnicity, education, marital_status, poverty_level_category, ever_received_food_stamps, worried_house_run_out_food, health_insurance, private_insurance, medicare, medicaid, food_ran_out) %>% 
  na.omit()

social_heart <- glm(heart_attack ~ gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid, data = social, family = "binomial")

social_stroke <- glm(stroke ~gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid, data = social, family = "binomial")

summary(step(social_heart, trace = 0))
summary(step(social_stroke, trace = 0))
```

### Health

```{r}
#health data
health <- data %>% 
  select(heart_attack, stroke, rfm, num_healthcare_visits, risk_sum) %>% 
  na.omit()

health_heart <- glm(heart_attack ~ rfm + num_healthcare_visits + risk_sum, data = health, family = "binomial")

health_stroke <- glm(stroke ~ rfm + num_healthcare_visits + risk_sum, data = health, family = "binomial")

summary(step(health_heart, trace=0))
summary(step(health_stroke, trace=0))
```

### Religious attendance

```{r}
#religious data
religion <- data %>% 
  select(heart_attack, stroke, religious_attendance) %>% 
  na.omit()

religion_heart <- glm(heart_attack ~ religious_attendance, data = religion, family = "binomial")

religion_stroke <- glm(stroke ~ religious_attendance, data = religion, family = "binomial")

summary(step(religion_heart, trace=0))
summary(step(religion_stroke, trace =0))
```

```{r, warn = FALSE, warning = FALSE}
#make tables
social_heart_tbl <- tbl_regression(social_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)  

social_stroke_tbl <- tbl_regression(social_stroke, exponentiate = TRUE) %>%
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

health_heart_tbl <- tbl_regression(health_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

health_stroke_tbl <- tbl_regression(health_stroke, exponentiate = TRUE) %>%
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

relig_heart_tbl <- tbl_regression(religion_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

relig_stroke_tbl <- tbl_regression(religion_stroke, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)
```

### Heart attack models

```{r}
tbl_merge(tbls = list(social_heart_tbl, health_heart_tbl, relig_heart_tbl),
          tab_spanner = c("Social Determinants", "Health and Behavior", "Religious Attendance"))
```

### Stroke models


```{r}
tbl_merge(tbls = list(social_stroke_tbl, health_stroke_tbl, relig_stroke_tbl),
          tab_spanner = c("Social Determinants", "Health and Behavior", "Religious Attendance"))
```

# Religious Attendance

```{r}
data_cat <- data %>% 
  select(-c(SEQN, age, born_outside_us, poverty_level_index, FSQ171,num_months_WIC, risk_mobility, risk_bp, risk_glucose, avg_drinks_daily_past_yr, num_drinks_past_yr,num_drinks_past_yr_unit, num_days_5_drinks_or_more, ALQ140U, had_12_drinks_lifetime, num_drinks_weekly,risk_alcohol, risk_tobacaco, religious_attendance_num, weight, height, waist, rfm, risk_sum, risk_rfm, heart_attack_num, stroke_num)) %>%  na.omit()

data.frame(Variables = colnames(data_cat),
           P = c(NA,
                 (chisq.test(data_cat$religious_attendance, data_cat$gender)[[3]]),
                 chisq.test(data_cat$religious_attendance, data_cat$ethnicity)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$education)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$marital_status)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$poverty_level_category)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$health_insurance)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$private_insurance)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$medicare)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$medicaid)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$num_healthcare_visits)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$ever_received_food_stamps)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$worried_house_run_out_food)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$food_ran_out)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$mobility_limit)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$high_blood_pressure)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$fasting_glucose)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$tobacco_use_5_days)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$had_12_drinks_in_1yr)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$type_of_drinker)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$religious_attendance)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$heart_attack)[[3]],
                 chisq.test(data_cat$religious_attendance, data_cat$stroke)[[3]]
                 )) %>% 
  mutate(`.05` = case_when(
    P <= 0.05 ~ "*",
    TRUE ~ as.character(".")
  )) %>% 
  kable() %>% 
  kable_styling("striped")
```

# Cumulative Models

### Social

```{r}
social_heart <- glm(heart_attack ~ gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid, data = data, family = "binomial")

summary(social_heart)


social_stroke <- glm(stroke ~gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid, data = data, family = "binomial")

summary(social_stroke)
```

### Social + Health

```{r}
social_health_heart <- glm(heart_attack ~ gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid + num_healthcare_visits + risk_sum, data = data, family = "binomial")

summary(social_heart)


social_health_stroke <- glm(stroke ~gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid + num_healthcare_visits + risk_sum, data = data, family = "binomial")

summary(social_stroke)
```

### Social + Health + Religion

```{r}
social_health_rel_heart <- glm(heart_attack ~ gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid + num_healthcare_visits + risk_sum + religious_attendance, data = data, family = "binomial")

summary(social_heart)


social_health_rel_stroke <- glm(stroke ~gender + age + ethnicity + education + marital_status
                     + poverty_level_category + ever_received_food_stamps + worried_house_run_out_food + food_ran_out + health_insurance + private_insurance + medicare + medicaid + num_healthcare_visits + risk_sum + religious_attendance, data = data, family = "binomial")

summary(social_stroke)
```

```{r}
#make tables
social_heart_tbl <- tbl_regression(social_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)  

social_stroke_tbl <- tbl_regression(social_stroke, exponentiate = TRUE) %>%
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

social_health_heart_tbl <- tbl_regression(social_health_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

social_health_stroke_tbl <- tbl_regression(social_health_stroke, exponentiate = TRUE) %>%
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

social_health_rel_heart_tbl <- tbl_regression(social_health_rel_heart, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)

social_health_rel_stroke_tbl <- tbl_regression(social_health_rel_stroke, exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Ref."
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)
```


### Heart Attack Cumulative Models

```{r}
tbl_merge(tbls = list(social_heart_tbl, social_health_heart_tbl, social_health_rel_heart_tbl),
          tab_spanner = c("Social Determinants", "Health and Behavior", "Religious Attendance"))
```

### Stroke Cumulative Models

```{r}
tbl_merge(tbls = list(social_stroke_tbl, social_health_stroke_tbl, social_health_rel_stroke_tbl),
          tab_spanner = c("Social Determinants", "Health and Behavior", "Religious Attendance"))
```

# Odds Ratio validation

```{r}
glm(heart_attack ~ ethnicity, "binomial", data = data) %>% 
  tbl_regression(exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Reference"
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)
```

```{r}
glm(heart_attack ~ relevel(ethnicity, ref = 3), "binomial", data = data) %>% 
  tbl_regression(exponentiate = TRUE) %>% 
  modify_table_styling(
    columns = c(estimate, ci),
    rows = reference_row %in% TRUE,
    missing_symbol = "Reference"
  ) %>% 
  modify_column_hide(column = ci) %>% 
  bold_labels() %>% 
  bold_p(t = 0.05)
```


# Age vs Cardiovascular Disease Correlation

```{r}
cor.test(data$heart_attack_num, data$age, method = c("pearson"))
```

```{r}
cor.test(data$stroke_num, data$age, method = c("pearson"))
```

